---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

#### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
# libraries
library(ggplot2)
library(dplyr)

if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```

#### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

## What is mean total number of steps taken per day?
```{r}
stepsTakenEachDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

#### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsTakenEachDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

#### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsTakenEachMean <- mean(stepsTakenEachDay)
stepsTakenEachMedian <- median(stepsTakenEachDay)
```
* Mean: `r stepsTakenEachMean`
* Median:  `r stepsTakenEachMedian`


## What is the average daily activity pattern?
```{r}
averageDailyActivityPattern <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

#### 1. Make a time series plot
```{r}
ggplot(data=averageDailyActivityPattern, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps <- which.max(averageDailyActivityPattern$meanSteps)
theMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageDailyActivityPattern[maxSteps,'interval'])
```

* Most Steps at: `r theMostSteps`


## Imputing missing values
#### 1. Calculate and report the total number of missing values in the dataset 
```{r}
missingValuesDataset <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r missingValuesDataset`

#### 2. Devise a strategy for filling in all of the missing values in the dataset.
#### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityDataImputing <- activityData
activityDataImputing$steps <- impute(activityData$steps, fun=mean)
```


#### 4. Make a histogram of the total number of steps taken each day 
```{r}
stepsByDayImputing <- tapply(activityDataImputing$steps, activityDataImputing$date, sum)
qplot(stepsByDayImputing, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

#### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsByDayMeanImputing <- mean(stepsByDayImputing)
stepsByDayMedianImputing <- median(stepsByDayImputing)
```
* Mean (Imputing): `r stepsByDayMeanImputing`
* Median (Imputing):  `r stepsByDayMedianImputing`


## Are there differences in activity patterns between weekdays and weekends?
#### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityDataImputing$dateType <-  ifelse(as.POSIXlt(activityDataImputing$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

#### 2. Make a panel plot containing a time series plot

```{r}
averagedActivityDataImputing <- aggregate(steps ~ interval + dateType, data=activityDataImputing, mean)
ggplot(averagedActivityDataImputing, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```




